---
title: "Hw-05-ChenchenGuo"
author: "Chenchen GUO"
date: "October 14, 2018"
output: github_document
always_allow_html: yes
---
-   [Introduction](#introduction)

    -[Goals](#goals)
-   [Part1: Factor management](#part-1-factor-management)

    -[Requirements](#requirements)
    
    -[Implementation](#implementation)
    
-   [Part2: File I/O](#part-2-file-io)

    -[Requirements2](#requirements2)
    
    -[Implementation2](#implementation2)
    
-   [Part3: Visualization design](#part-3-visualization-design)
    
    -[Requirements3](#requirements3)
    
    -[Implementation3](#implementation3)

-   [Part4: Writing figures to file](#part-4-writing-figures-to-file)
    
    -[Requirements4](#requirements4)
    
    -[Implementation4](#implementation4)

Introduction
------------

[Homework 05](http://stat545.com/Classroom/assignments/hw05/hw05.html): Factor and figure management
The aim of the homework is to have a better understanding of factor and figure management, also including the file I/O implementation and manipulating
data. In this implementation, package gapminder is used.

Goals:
-----------

1. Reorder a factor in a principled way based on the data and demonstrate the effect in arranged data and in figures.

2. Write some data to file and load it back into R.

3. Improve a figure (or make one from scratch), using new knowledge, e.g., control the color scheme, use factor levels, smoother mechanics.

4. Make a plotly visual.

5. Implement visulization design principles.


Part 1: Factor management
-------------------------

Requirements
------------
1. Drop factor/levels
2. Recorder levels based on knowledge from data
3. Explore the effects of arrange()

Implementation
--------------

Firstly load all needed packages

```{r}
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(scales))
```
1. List the gapminder continent information before drop oceania

```{r}
con_whole <- gapminder$continent
levels(con_whole)
#fct_count(con_whole)
#Form a table to show original continent information
knitr::kable(fct_count(con_whole), col.names = c('Continent', 'Number of countries')) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)
```
Then use drop()
```{r}
con_drop <- gapminder %>% 
  filter(continent != 'Oceania') %>% 
  droplevels()
#Indicate the number of continents after drop
levels(con_drop$continent)
#Show same continents without oceania
knitr::kable(fct_count(con_drop$continent), col.names = c('Continent', 'Number of countries')) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)
```

2. Reorder levels 
Here I will use the lifeexpectancy as a standard to level different continents.
For example, use the variance of lifeExp 
```{r}
#Use summarize function to calculate all variance of lifeExp for all continents
var_life <- gapminder %>% 
  group_by(continent) %>% 
  summarise(variance = var(lifeExp))
#Form a table with variance and continent
knitr::kable(var_life, col.names = c('Continent', 'Variance of life expectancy')) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)



```

```{r}
#Visulization
lifescales <- ggplot(var_life, aes(continent, variance))+
               geom_col()+
               ggtitle("Variance of life expectancy of continent")
# In order to evidently showing the number, add a scales for Y-axis
lifescales +
  scale_y_continuous(breaks = 1: 20 * 10)
```


Then use fct_reorder() to order continents by variance of lifeExp
```{r}
life_ord <- mutate(var_life, continent = fct_reorder(continent, variance))
# Visualization , this time the order of data changed from the above
ggplot(life_ord, aes(variance, continent))+
  geom_point()+
  scale_x_continuous(breaks = 1:15*10)+
  ggtitle("Variance of life expectancy of continent")
```

Now show the difference between fct_reorder() and arrange()

```{r}
# for fct_reorder list the table:
knitr::kable(life_ord, col.names = c('Continent', 'Variance of life expectancy')) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

```
The data sequence in table are still the same, now use arrange()
```{r}
arr_life <- var_life %>% 
  arrange(variance)
knitr::kable(arr_life, col.names = c('Continent', 'Variance of life expectancy')) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

# Now the table of continent vs. var of life  changed order
ggplot(arr_life, aes(variance, continent))+
  geom_point()
#Here the arrange function only alter the sequence in table but not in figure
```

Hence, if we want to both alter the sequence on figure and table
```{r}
final_life <- var_life %>% 
  mutate(continent = fct_reorder(continent, variance)) %>% 
  arrange(variance)
# visulization of both table and figure:
knitr::kable(final_life, col.names = c('Continent', 'Variance of life expectancy')) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

ggplot(final_life, aes(variance, continent))+
  geom_point()+
  scale_x_continuous(breaks = 1:15*10)+
  ggtitle("Variance of life expectancy of continent")

#Now both table and figure's order in terms of variance of lifeexp are increasing
```





Part 2: File I/O
----------------

Requirements2
------------
 Experiments with one or more of write_csv()/read_csv(), saveRDS()/readRDS(), dput()/dget().
 Make them non-alphabetical 


Implementation2
--------------
```{r}

# Firstly, use the write_csv() and read_csv() function
write.csv(arr_life, "Life_arrange.csv", append = FALSE, col.names = TRUE)
#Then to read from this csv file

a <- read.csv("Life_arrange.csv")
# To visualize csv file
knitr::kable(a) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

#the str() function indicates the difference between original data frame and after writing into csv
# After writing into .csv file , there is one extra sequence column 
str(arr_life)
str(a)

```
Now use saveRDS() / readRDS() function

```{r}
saveRDS(arr_life, "Life_arrange.rds")
b <- readRDS("Life_arrange.rds")

# To visualize rds file
knitr::kable(b) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

# now we see the difference between original data and after writing into .rds file
# Here, the saveRDS() function keeps the original data.
str(arr_life)
str(b)

```

Now use the dput()/ dget()
```{r}
# same as before, use the arr_life data frame
dput(arr_life, "Life_arrange.R")
c <- dget("Life_arrange.R")

# To visualize the .R file
knitr::kable(c) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

# Here, same as the saveRDS() function, the dput() function also keeps the original data.
str(arr_life)
str(c)
```


Part 3: Visualization design
-----------------------------

Requirements3
------------
 Use the country or continent color scheme that ships with Gapminder.
 Use plotly()
 compare with original ggplot


Implementation3
--------------
```{r}
# Firstly, the original ggplot of gapminder of different continents showing lifeExp vs. gdpPercap
p <- gapminder %>% 
  filter(continent != "Oreania") 
str(p)
ggplot(p, aes(gdpPercap, lifeExp)) + 
  geom_point()+
  facet_wrap( ~ continent)

# This is the original figure, now we add some labels, scales and colour scheme
ggplot(p, aes(gdpPercap, lifeExp)) +
  geom_point(aes(colour = pop), alpha = 0.5) +
  scale_x_log10(labels = dollar_format()) +
  scale_color_distiller(
    trans = "log10",
    breaks = 10^(1:10),
    labels = comma_format(),
    palette = "Blues"
  )+
  scale_y_continuous(breaks = 10 * (1:10))+
  facet_wrap( ~ continent, scales = "free_y")+
   theme_bw()+
  theme(
    axis.text = element_text(size = 10),
    strip.background = element_rect(fill = "light grey"),
    panel.background = element_rect(fill = "white")
        )+
  ggtitle("Life expectation VS. GDP")
  

  
```

```{r}
# Another example
# Orignal figure
gapminder %>% 
  ggplot(aes(gdpPercap, fill=continent))+
  facet_wrap(~continent)+
  geom_density()+xlab("GDPpercap")+ylab("Frequency")+
  ggtitle("GDP distribution of each continent")


# After design
gapminder %>% 
  ggplot(aes(gdpPercap, fill=continent)) +
  geom_density(aes(colour = pop), alpha = 0.2)+
  scale_x_log10(labels = dollar_format()) +
    scale_color_distiller(
    trans = "log10",
    breaks = 10^(1:10),
    labels = comma_format(),
    palette = "Blues"
  ) +
  facet_wrap( ~ continent)+
  ggtitle("GDP distribution")+
  theme_bw()+
  theme(
    axis.text = element_text(size = 10),
    strip.background = element_rect(fill = "light blue"),
    panel.background = element_rect(fill = "white")
        )
```


For the implementation of ploty() function

```{r, warning=FALSE}
# library(plotly)
# (h <- gapminder %>% 
#   filter(continent != "Oceania") %>% 
#   ggplot(aes(gdpPercap, lifeExp))+
#   geom_point(aes(colour = pop), alpha = 0.5) +
#   scale_x_log10(labels = dollar_format()) +
#   scale_color_distiller(
#     trans = "log10",
#     breaks = 10^(1:10),
#     labels = comma_format(),
#     palette = "Blues"
#   )+
#   scale_y_continuous(breaks = 10 * (1:10))+
#   facet_wrap( ~ continent, scales = "free_y")+
#    theme_bw()+
#   theme(
#     axis.text = element_text(size = 10),
#     strip.background = element_rect(fill = "light grey"),
#     panel.background = element_rect(fill = "white")
#         )+
#   ggtitle("Life expectation VS. GDP"))
#   
# n<- ggplotly(h)
# api_create(n, filename = "r-docs-ggplotly")


```

[Ggplotly](https://plot.ly/~GChenchen/5/#/)

```{r, warning=FALSE}
# m<- plot_ly(gapminder,
#         x = ~gdpPercap, 
#         y = ~lifeExp, 
#         z = ~pop,
#         type = "scatter3d",
#         mode = "markers",
#         opacity = 0.2) %>% 
#   layout(
#     scene = list(
#       xaxis = list(title = "gdpPercap"),
#       yaxis = list(title = "LifeExp"),
#       zaxis = list(title = "Population")))
# 
# api_create(m, filename = "r-docs-plotly")
```

[plotly_3D1](https://plot.ly/~GChenchen/3/#/)

3D scattering
```{r, warning=FALSE}
# o<-plot_ly(gapminder, 
#         x = ~gdpPercap, 
#         y = ~year, 
#         z = ~pop,
#         type = "scatter3d",
#         mode = "markers",
#         opacity = 0.2) %>% 
#   layout(xaxis = list(type = "log"),
#           scene = list(
#       xaxis = list(title = "gdpPercap"),
#       yaxis = list(title = "Year"),
#       zaxis = list(title = "Population")
#          ))
# 
# api_create(o, filename = "r-docs-plotly 3d2")
```

[Plotly_3D2](https://plot.ly/~GChenchen/7/#/)

Part 4: Writing figures to file
-------------------------------

Requirements4
------------
 Use ggsave() to explicitly save a plot to file. Then load and embed in report.
 Play the arguments of ggsave(), such as width, height, resolution or text scaling
 Various graphics devices: a vector vs. raster format
 Explicit provision of the plot object p via ggsave(..., plot = p). showa a situation in which this actually matters.
 
 
Implementation4
--------------
Use ggsave() to save above figures separately in .png and .pdf format
```{r}
figure1 <-gapminder %>% 
  ggplot(aes(gdpPercap, fill=continent)) +
  geom_density(aes(colour = pop), alpha = 0.2)+
  scale_x_log10(labels = dollar_format()) +
    scale_color_distiller(
    trans = "log10",
    breaks = 10^(1:10),
    labels = comma_format(),
    palette = "Blues"
  ) +
  facet_wrap( ~ continent)+
  ggtitle("GDP distribution")+
  theme_bw()+
  theme(
    axis.text = element_text(size = 10),
    strip.background = element_rect(fill = "light blue"),
    panel.background = element_rect(fill = "white")
        )
  
ggsave("figure1.png",
       plot = figure1,
       width = 15,
       height = 10,
       scale = 1.5)

figure2 <- gapminder %>% 
  filter(continent != "Oreania") %>% 
  ggplot(aes(gdpPercap, lifeExp)) +
  geom_point(aes(colour = pop), alpha = 0.5) +
  scale_x_log10(labels = dollar_format()) +
  scale_color_distiller(
    trans = "log10",
    breaks = 10^(1:10),
    labels = comma_format(),
    palette = "Blues"
  )+
  scale_y_continuous(breaks = 10 * (1:10))+
  facet_wrap( ~ continent, scales = "free_y")+
   theme_bw()+
  theme(
    axis.text = element_text(size = 10),
    strip.background = element_rect(fill = "light grey"),
    panel.background = element_rect(fill = "white")
        )+
  ggtitle("Life expectation VS. GDP")


ggsave("figure2.pdf",
       plot = figure2,
       width = 15,
       height = 10,
       scale = 1.5)
```









